Fusion Moves for Correlation Clustering

### 2014-11-12 paper/egpaper_for_review.tex

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning, \cf Fig.~\ref{fig:fusion_abiguty}.
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}_{ij}    & = \max\{ y_{ij}', y_{ij}''\}  & \forall {ij}\in E\\  % y' \cup y'' \\
\mathcal{E}_0  & =  \{ (i, j) \; | \; \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{E}_0$.
%
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
      &\quad \textrm{s.t.} \quad y_{ij} = 0 & \forall (i, j) \in \mathcal{E}_0 \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy,
because $y'$ and $y''$ are feasible solutions for problem~\ref{eq:fusion_move_a}.
%\begin{align}
%  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
%\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solution.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.
%
How to optimize  Eq.~\ref{eq:fusion_move_a} in an more efficient way is
described in the following section.

@@

\subsection{Fast Optimization} \label{sec:fast_optimization}

Eq.~\ref{eq:fusion_move_a} can be solved efficiently, since 
we can work on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes.
This coarser graph is constructed by contracting all must-link edges
(see Fig.~\ref{fig:move_graph}).

\begin{definition}\emph{(Contradicted Graph)}
Given a weighted graph $G=V,E,w)$ segmentation of $G$ given by $y\in P(G)$,
we denote by $S: V \to \mathbb{N}$ a mapping from the nodes to the segments.
We define the contradiction of graph $G_y=(V_y,E_y, \bar{w})$
by $V_y=\{S(v)|v\in V\}$, $E_y=\{S(u)S(v)|uv\in E\}$, and 
$\forall \bar{u}\bar{v}\in E_y: \bar{w}_{\bar{u}\bar{v}}=\sum_{uv\in E, S(u)=\bar{u},S(v)=\bar{v}}w_{uv}$ 
\end{definition}

\begin{definition}\emph{(Back Projection)}
Given a segmentation $\bar{y}$ of a contradicted graph $G_y=(V_y,E_y)$.
We define the back projection $\rho^G$ from $\bar{y}$ to a segmentation of original graph  $G=(V,E)$
by 
\begin{align}
\rho^G_{uv}(\bar{y})=\left\{ 
\begin{array}{ll}
\bar{y}_{S(u)S(v)}& \textrm{ if }S(u) \neq S(v)\\
0                & \textrm{ else}
\end{array}
\right.
\end{align}
\end{definition}

\begin{theorem}[Equivalence]
The optimal multicut $\bar{y}^*$ on the contracted graph $G_y$
is equivalent to the optimal multicut $y^*$ in the original graph $G$ 
with additional must link constraints ($y^* \leq y$).
%
\begin{proof}
Since $y^*$ fulfills all  must link constraints, we have  $y^*\cdot w = \rho^{-1}_y(y^*)\cdot \bar{w}$.
Consequently, $y^*\cdot w = \rho_y(\bar{y}) \cdot w$ and  $\rho^{-1}_y(y^*)\cdot \bar{w}= \bar{y}\cdot\bar{w}$.
\end{proof}
%
\end{theorem}

TODO:
Here we need to proof that it is equivalent 
to optimize the contracted graph !

\input{inputs/fig_contraction.tex}
\input{inputs/fig_projection.tex}

>>>>

Fusion moves as defined in \cite{Lempitsky-2010} work
in the node domain and do not work properly for 
objective functions as Eq.~\ref{eq:nodeproblem} since
the node coloring is ambiguous and has no semantic meaning, \cf Fig.~\ref{fig:fusion_abiguty}.
In the following, we propose a more suitable fusion move for correlation
clustering which works on the edge domain.
%
Given two proposal solutions $y'$ and $y''$,
$\mathcal{C}$ is the set of edges
which are uncut in $y'$ and $y''$.
%
\begin{align}
\breve{y}_{ij}    & = \max\{ y_{ij}', y_{ij}''\}  & \forall {ij}\in E\\  % y' \cup y'' \\
\mathcal{E}_0^{\breve{y}}  & =  \{ ij \in E \; | \; \breve{y}_{ij} = 0 \}
\end{align}
%
The fusion move for correlation clustering is solving Eq.~\ref{eq:edgeproblem}
with additional \emph{must-link constraints} for all edges in $\mathcal{E}_0^{\breve{y}}$.
%
\begin{align}
  y^* &= \argmin_{y \in P(G)} \sum_{ (i,j) \in E } w_{ij} \cdot y_{ij} \label{eq:fusion_move_a}.\\ 
      &\quad \textrm{s.t.} \quad y_{ij} = 0 & \forall (i, j) \in \mathcal{E}_0^{\breve{y}} \nonumber
\end{align}
%
By construction, solving Eq.~\ref{eq:fusion_move_a} cannot increase the energy,
because $y'$ and $y''$ are feasible solutions for problem~\ref{eq:fusion_move_a}.
%\begin{align}
%  \sum  w_{ij} \cdot y^*_{ij}  \leq \min \left( \sum  w_{ij} \cdot y'_{ij} ,  \sum   w_{ij} \cdot y''_{ij} \right)
%\end{align}
%

Once we know how to fuse segmentations, the actual algorithm 
for correlation clustering 
is simple and similar to the fusion move algorithm of Lempitsky \etal~\cite{bagon_2011_arxiv}.
We iteratively improve the best solution by
fusing it with proposal solution.
The inherent difference is how we define the fusion.
We fuse the segmentations according to Eq.~\ref{eq:fusion_move_a}.

A further difference is how we efferently calculate the correlation clustering fusion move and how we generate proposals.
Both will be discussed next. The overall framework is sketched in Fig.~\ref{fig:move_graph}.

@@

\subsection{Fast Optimization of CCFusion Moves} \label{sec:fast_optimization}
In general the auxiliary fusion problem~\ref{eq:fusion_move_a} is still NP-hard.
However, many variables have been fixed to be zero and we can reformulate~\ref{eq:fusion_move_a}
into a correlation clustering problem on a coarser graph, where all nodes which are connected
via must-link constraints are merged into single nodes. We call this graph \emph{contracted graph}.
%
\begin{definition}\emph{(Contracted Graph)}
Given a weighted graph $G=V,E,w)$ segmentation of $G$ given by $y\in P(G)$,
we denote by $S: V \to \mathbb{N}$ a mapping from the nodes to the segments.
We define the contraction of graph $G_y=(V_y,E_y, \bar{w})$
by $V_y=\{S(v)|v\in V\}$, $E_y=\{S(u)S(v)|uv\in E\}$, and 
$\forall \bar{u}\bar{v}\in E_y: \bar{w}_{\bar{u}\bar{v}}=\sum_{uv\in E, S(u)=\bar{u},S(v)=\bar{v}}w_{uv}$ 
\end{definition}
%
Any clustering $\bar{y}$ of the contracted graph $G_y=(V_y,E_y)$ can be \emph{back projected} to a clustering $y$ the original graph $G=(V,E)$ by
%
\begin{align}
y_{uv}&=\left\{ 
\begin{array}{ll}
\bar{y}_{S(u)S(v)}& \textrm{ if }S(u) \neq S(v)\\
0                & \textrm{ else}
\end{array}
\right.& \forall uv \in E
\end{align}
Furthermore, the back projection of the optimal segmentation $\bar{y}$ of the contracted graph is a solution of problem~\ref{eq:fusion_move_a}.

% Eq.~\ref{eq:fusion_move_a} can be solved efficiently, since 
% we can work on a coarser graph, where all nodes which are connected
% via must-link constraints are merged into single nodes.
% This coarser graph is constructed by contracting all must-link edges
% (see Fig.~\ref{fig:move_graph}).

% \begin{definition}\emph{(Back Projection)}
% Given a segmentation $\bar{y}$ of a contradicted graph $G_y=(V_y,E_y)$.
% We define the back projection $\rho^G$ from $\bar{y}$ to a segmentation of original graph  $G=(V,E)$
% by 
% \begin{align}
% \rho^G_{uv}(\bar{y})=\left\{ 
% \begin{array}{ll}
% \bar{y}_{S(u)S(v)}& \textrm{ if }S(u) \neq S(v)\\
% 0                & \textrm{ else}
% \end{array}
% \right.
% \end{align}
% \end{definition}

\begin{theorem}[Equivalence]
The back projection of the optimal segmentation $\bar{y}'$ of the contracted graph $G_y=(V_y,E_y, \bar{w})$
is an optimal solution of problem~\ref{eq:fusion_move_a}.
\begin{proof}
Let $y'$ be the back propagation of $\bar{y}'$, which is by definition a feasible for~\ref{eq:fusion_move_a}.
If $y'$ would not be an optimal solution than there must exists $y''$ with 
$\sum_{e\in E} w_ey_e'' > \sum_{e\in E} w_ey_e''$.
Since $y_e'$ and $y_e''$ are $0$ for all $e \in \mathcal{E}_0^{\breve{y}}$ we would have
\begin{align*}
  \sum_{\bar{e}\in E_y} \bar{w}_{\bar{e}}\bar{y}'_{\bar{e}}
 &= \sum_{e\in E \setminus \mathcal{E}_0^{\breve{y}}} w_ey_e'          
  = \sum_{e\in E} w_ey_e'' \\
 &> \sum_{e\in E} w_ey_e''
  = \sum_{e\in E \setminus \mathcal{E}_0^{\breve{y}}} w_ey_e''
  =\sum_{\bar{e}\in E_y} \bar{w}_{\bar{e}}\bar{y}''_{\bar{e}}
\end{align*}
where $\bar{y}''$ is the projection from $y''$ on $G_y$.
This contradicts that $y'$ is a optimal segmentation of $G_y$.
\end{proof}
%
\end{theorem}

Instead of problem~\ref{eq:fusion_move_a} we can now solve problem~ref{eq:edgeproblem} on the contracted graph $G_{\breve{y}}$.
This is, depending on the intersection of the current and proposed solution, magnitudes smaller than $G$.
The correlation clustering problem on $G_{\breve{y}}$ can be solved by any correlation clustering solver.
Since $G_{\breve{y}}$ is smaller exact methods or good approximative methods like 
multicuts~\cite{kappes_2013_arxiv} or CGC~\cite{beier_2014_cvpr}
become tractable.

%\input{inputs/fig_contraction.tex}
%\input{inputs/fig_projection.tex}
